Boutiques: a flexible framework for automated application integration in computing platforms

### 2017-10-12 boutiquespaper.tex
updated methods section

At runtime, a value is assigned to all the mandatory and some of the
optional inputs. %% Inputs of type ``String'' may contain any string,
%% inputs of type "Number" must contain a string representing a number,
%% inputs of type "File" must contain a string representing a file path
%% (absolute or relative to the execution directory) and inputs of type
%% "Flag" must contain a boolean. 
%% When input is a list, the value contains the concatenation of all the
%%values in the list.
Algorithm~\ref{algo:command-line} shows how the command line is
constructed from the manifest and the values entered by the user. It
substitutes all the value keys in the command line, output path
templates and configuration files, and writes the configuration files.
\begin{algorithm}[h!]
\caption{Command-line construction}
\label{algo:command-line}
\begin{algorithmic}
  \State \Comment{Substitute input value keys in output path templates, configuration files and command line.}
  \For{\texttt{input} in \texttt{inputs}}
  \If{\texttt{input} has a \texttt{value-key}}
  \For{\texttt{output} in \texttt{outputs}}
  \State \texttt{stripped\_value} = \texttt{input\_value}
  \If{\texttt{input} type is \texttt{File}}
  \State In \texttt{stripped\_value}, remove  all elements in \texttt{path-template-stripped-extensions}.
  \EndIf
  \State In \texttt{path-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \If{\texttt{output} has a \texttt{file-template}}
  \State In any line of \texttt{file-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \EndIf
  \EndFor
  \State Prepend \texttt{command-line-flag} and \texttt{command-line-flag-separator} to \texttt{input\_value}.
  \State In \texttt{command-line}, replace all occurrences of \texttt{value-key} by \texttt{input\_value}.
  \EndIf
  \EndFor

@@

Invocation schemas, however, are complex JSON objects. Basically, they
must represent the properties described above in a formal way,
including dependencies between
inputs. Listing~\ref{listing:invocation-schema-example} shows an
example of how dependencies between mutually exclusive parameters are
defined in the invocation schema. As it does not seem realistic to
assume that developers will easily write correct invocations schemas
even for their own applications, invocation schemas can be generated
automatically using a core \boutiques tool. The invocation schema is
stored as an optional property of the Boutiques manifest.

@@

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
"dependencies" : {
      "num_input" : {
         "properties" : {
            "str_input" : {
               "not" : {}
            }
         }
      },
      "str_input" : {
         "properties" : {
            "num_input" : {
               "not" : {}
            }
         }
      }
   }
\end{minted}
\caption{Excerpt from invocation schema showing dependencies between
  two mutually exclusive parameters \texttt{num\_input} and
  \texttt{str\_input}.}
\label{listing:invocation-schema-example}
\end{listing}

@@

Applications are installed in a container image complying to the
Docker, Singularity or rootfs format. We intentionally support
multiple container formats as we anticipate that they will be used for
different purposes. For instance, Docker is well suited for developers
and users who manipulate applications on their local workstation. It
is well documented, maintained and it has a rich ecosystem of tools to
help build and run containers on most operating systems. Singularity
is more suited for users and platforms that need to run applications
on shared computing clusters. Bridges exist among these containers
formats to convert container images across frameworks. For instance, a
platform dedicated to high-performance computing may accept manifests
referring to Docker containers to facilitate application integration
by developers, and internally convert container images to Singularity
to run applications efficiently on clusters.

Container images are defined from their URL (Singularity and rootfs)
or name in a Docker index. Container images may specify a working
directory where the application has to be run. A hash may also be
reported to accurately identify container images and detect updates.

@@

Containers are not mandatory though. If an application manifest does
not contain any reference to a container, the platform will assume
that the application is pre-installed on the execution infrastructure.

@@

Custom properties may be added to the Boutiques specification without
restriction. Custom properties are corralled in a specific JSON object
to facilitate validation. They may be useful to implement
platform-specific features but they should be used with care to avoid
making applications dependent on a particular platform.

@@

The Boutiques validator checks conformance of JSON manifests to the
Boutiques schema using a basic JSON validator. It also performs the
following checks that cannot be easily implemented in JSON schema:
value keys are unique among inputs, input and output identifiers are
unique, input and output value keys are all included in the command
line, value keys are not contained within each other (which would
puzzle substitution), output path templates are unique (to avoid
results to overwrite each other), inputs of type Flag have a
command-line flag, are optional and are not lists, number and list
constraints are sensible (e.g. min is lower than max), the default
value of restricted types is part of the restriction, an input cannot both require
and disable another input, required inputs cannot require or disable
other parameters, group member identifiers must correspond to existing
inputs and cannot appear in different groups, mutually exclusive
groups cannot have members requiring other members, and
one-is-required groups should never have required members. The
Boutiques validator is meant for application developers; platforms
should rely on regular JSON validation to avoid depending on a
third-party tool.

The local executor can generate and execute command line from a
Boutiques manifest and a set of input values represented in a CSV or
JSON file complying with the invocation schema. It runs the command in
a container provided that the required framework (e.g. Docker) is
already installed. It can also generate hypothetical command lines
from random values for debugging purposes. The local executor can be
used by application users to run applications locally, or by platforms
to generate command lines to be run on the execution infrastructure.

The invocation schema generator can create an invocation schema (see
Section~\ref{sec:invocation-schema}) from a \boutiques manifest and
validate input data against it using a regular JSON validator. It can
be used to add invocation schemas to existing manifests.

Finally, a JSON editor facilitates the writing of \boutiques manifests
through a web interface and a Jupyter Notebook tutorial facilitates
new users getting started with Boutiques.

\subsection{Repository}

\todo{Gunnar, Nathaniel, John}

@@

The following platforms are able to integrate and execute Boutiques
applications.

>>>>

At runtime, a value is assigned to all the mandatory and some of the
optional inputs. %% Inputs of type ``String'' may contain any string,
%% inputs of type "Number" must contain a string representing a number,
%% inputs of type "File" must contain a string representing a file path
%% (absolute or relative to the execution directory) and inputs of type
%% "Flag" must contain a boolean. 
%% When input is a list, the value contains the concatenation of all the
%%values in the list.
% gk: I agree with not including the above
Algorithm~\ref{algo:command-line} shows how the command line is
constructed from the manifest and the values entered by the user. It
substitutes all the value keys in the command line, output path
templates and configuration files, and writes the configuration files.
\begin{algorithm}[h!]
\caption{Command-line construction}
\label{algo:command-line}
\begin{algorithmic}
  \State \Comment{Substitute input value keys in output path templates, configuration files and command line.}
  \For{\texttt{input} in \texttt{inputs}}
  \If{\texttt{input} has a \texttt{value-key}}
  \For{\texttt{output} in \texttt{outputs}}
  \State \texttt{stripped\_value} = \texttt{input\_value}
  \If{\texttt{input} type is \texttt{File}}
  \State In \texttt{stripped\_value}, remove  all elements in \texttt{path-template-stripped-extensions}.
  \EndIf
  \State In \texttt{path-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \If{\texttt{output} has a \texttt{file-template}}
  \State In any line of \texttt{file-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \EndIf
  \EndFor
  \State Prepend \texttt{command-line-flag} and \texttt{command-line-flag-separator} to \texttt{input\_value}.
  \State In \texttt{command-line}, replace all occurrences of \texttt{value-key} by \texttt{input\_value}.
  \EndIf
  \EndFor

@@

Invocation schemas, however, are complex JSON objects. Basically, they
must represent the properties described above in a formal way,
including dependencies between
inputs. Listing~\ref{listing:invocation-schema-example} shows an
example of how dependencies between mutually exclusive parameters are
defined in the invocation schema. As it does not seem realistic to
assume that developers will easily write correct invocations schemas
even for their own applications, invocation schemas can be generated
automatically by the \boutiques command-line utility. The invocation
schema is stored as an optional property of the \boutiques manifest.

@@

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
  "dependencies" : {
      "num_input" : {
         "properties" : {
            "str_input" : {
               "not" : {}
            }
         }
      },
      "str_input" : {
         "properties" : {
            "num_input" : {
               "not" : {}
            }
         }
      }
   }
}
\end{minted}
\caption{Excerpt from invocation schema showing dependencies between
  two mutually exclusive parameters \texttt{num\_input} and
  \texttt{str\_input}.}
\label{listing:invocation-schema-example}
\end{listing}

@@

Applications are installed in a container image complying to the
Docker, Singularity or rootfs format. We intentionally support
multiple container formats as we anticipate that they will be used for
different purposes. For instance, Docker is well suited for developers
and users who manipulate applications on their local workstations or the
cloud. It is well documented, maintained and it has a rich ecosystem of
tools to help build and run containers on most operating systems. Singularity
is more suited for users and platforms that need to run applications
on shared computing clusters. Bridges exist among these containers
formats to convert container images across frameworks. For instance, a
platform dedicated to high-performance computing may accept manifests
referring to Docker containers to facilitate application integration
by developers, and internally convert container images to Singularity
to run applications efficiently on clusters.

Container images are defined from their URL (rootfs) or name in the
Docker or Singularity indices. Manifests and container images may specify a
working directory where the application has to be run, and a hash may also be
reported to accurately identify container images and detect updates.

@@

%%Containers are not mandatory though. If an application manifest does
%%not contain any reference to a container, the platform will assume
%%that the application is pre-installed on the execution infrastructure. 
% gk: I believe we only support containers now, correct?

@@

Custom properties may be added to the Boutiques specification without
restriction. Custom properties are corralled in a specific JSON object
to facilitate validation. They may be useful to implement
platform-specific features but they should be used with care to avoid
making applications dependent on a particular platform or replacing
existing functionality already represented in the \boutiques schema.

@@

\boutiques is available on Python through the PyPi package repository as
``\texttt{boutiques}''. The \boutiques package exposes a command-line
utility to the user, \texttt{bosh}, which contains entrypoints for all
core functions within \boutiques. There is help-text which should guide
users through interacting with the \texttt{bosh} utility. The core tools
provided by \boutiques are, the: validator, executor (for both launch and
simulation), invocation schema handler, importer, and publisher. The tools
exposed through the command-line API are also available consistently through
the Python interface by importing the \boutiques package.

\paragraph{Validator.} The \boutiques validator checks conformance of JSON
manifests to the \boutiques schema using a basic JSON validator. It also
performs the following checks that cannot be easily implemented in JSON schema:
value keys are unique among inputs, input and output identifiers are
unique, input and output value keys are all included in the command
line, identifiers with the same value key are mutually exclusive, value keys
are not contained within each other (which would puzzle substitution), output
path templates are unique (to avoid results to overwrite each other), inputs of
type Flag have a command-line flag, are optional and are not lists, the default
value of restricted types is part of the restriction, an input cannot both require
and disable another input, required inputs cannot require or disable
other parameters, group member identifiers must correspond to existing
inputs and cannot appear in different groups, mutually exclusive
groups cannot have members requiring other members, one-is-required groups
should never have required members, and all-or-none group members must not be
required.

\paragraph{Executor.} The executor has two modes of operation: \emph{launch}, and
\emph{simulate}. The simulate mode can generate hypothetical command lines from
random values given the manifest (and corresponding invocation schema) for debugging
purposes, or display the command that would be executed given a provided valid
invocation. The launch mode can execute command line from a \boutiques manifest and
a set of input values represented in JSON file complying with the invocation schema.
It runs the command in a container provided that the required framework (e.g. Docker)
is already installed. The executor can be used by application users to run
applications locally, or by platforms to generate command lines to be run on the
execution infrastructure.

\paragraph{Invocation Schema Handler.} The invocation schema handler can create an
invocation schema (see Section~\ref{sec:invocation-schema}) from a \boutiques manifest
and validate input data against it using a regular JSON validator. It can be used to add
invocation schemas to existing manifests. It is used by the executor if no invocation
schema is present in the \boutiques manifest being deployed.

\paragraph{Importer.} The importer takes \boutiques manifests from older versions and
updates them to be compliant with the most recent version of the schema. This tool can
also create manifests from select application collections, such as BIDS
apps~\cite{gorgolewski2017bids}.

\paragraph{Publisher.} As \boutiques has primarily been adopted in the neuroinformatics
community, the publisher gets a further description (such as author, website, etc.) of
the described tool, and adds an index to it on
NeuroLinks\footnote{\url{https://brainhack101.github.io/neurolinks/}}, a repository
containing links to useful resources and tools. This enables others to find and use the
descriptor for your tool easily. This functionality could be extended to
new repositories for other domains, such as Bioconductor for bioinformatics.

Though not a component of the \boutiques tool chain, a Jupyter Notebook tutorial exists
to facilitate new users getting started with \boutiques.

\subsection{Repository}
%% gk: I think we should remove this and just have the reference in Publisher
\todo{Gunnar, Nathaniel, John}

@@

The following platforms are currently able to integrate and execute Boutiques
applications.
